# Copenhagen region {#study-area}

We focus the study on Copenhagen (*muni_id = 0101*) and Frederiksberg (*muni_id = 0147*) municipalities (Figure \@ref(fig:fig-CPH-region)A), and we make the analysis at a scale of grid cells of 100m x 100m (Figure \@ref(fig:fig-CPH-region)B). 

```{r adm-units, fig.cap="Parishes in the study area"}

# Communes polygons of Denmark, and select those in the study area
  dk_country <- st_union(dk_muni)
  
# Codes of the communes under study (KOMKODE)
  study_area_codes <- c("0101", "0147")

# Select municipalities under study  
  cph_muni <- dk_muni %>% filter(muni_id %in% study_area_codes) 
  
# Parishes polygons of Denmark, and select those in the study area 
  parish_link <- paste(loc_dir,
                       "DAGIREF_SHAPE_UTM32-EUREF89/ADM",
                       "SOGN.shp",
                       sep = "/")
  dk_parish  <- read_sf(parish_link) %>%
    st_zm() %>% 
    st_transform(crs = "EPSG:25832")

# Select those where the centroid is in the study area
  dk_parish_cent <- st_centroid(dk_parish)
  cph_parish_cent <- st_intersection(dk_parish_cent, cph_muni)

  cph_parish <- filter(dk_parish, SOGNEKODE %in% cph_parish_cent$SOGNEKODE) %>% 
    # Combine several parish features geometries into one polygon
    group_by(SOGNEKODE, SOGNENAVN) %>% 
    summarise(geometry = st_union(geometry)) %>% 
    ungroup() %>% 
    # add area of the parish (in km2)
    mutate(prsh_area_km2 = as.numeric(units::set_units(st_area(.), km^2)))

# Contour of the study area (merge the parishes in one polygon):
  study_area <- cph_parish %>%
    st_union() %>%
    st_sf() %>% 
    st_transform(crs = "EPSG:25832")

# Make grid cells of 100m x 100m 
grids100m <- study_area %>% 
  # Make regular grids (100m x 100m)
  st_make_grid(cellsize = 100) %>% 
  st_sf() %>% 
  # Select grids only in the study area
  mutate(int = st_intersects(., study_area) %>% lengths > 0) %>% 
  filter(int == TRUE) %>%
  select(-int) %>% 
  # Name grids as "g001", "g002", ...
  mutate(grid_ID = paste0("g", stringr::str_pad(seq(1, nrow(.), 1), 3, pad = "0")))

# Centroids of the grid cells (for Potential Model)  
grids100m <- grids100m %>%
  # Nesting
  nest(data_poly = everything()) %>% 
  # add centroids of the grids
  mutate(data_points = map(data_poly, st_centroid))

```

```{r fig-CPH-region, fig.width=10, fig.cap="Copenhagen and Frederiksberg municipalities"}

# Plot parish in the study area
p1 <- ggplot() + 
  geom_sf(data = cph_muni, fill = "grey", color = "grey50", size = 0.05) +
  my_theme_map() +
  labs(title = "A) Municipalities") 

# Plot grids 
p2 <- ggplot() + 
  geom_sf(data = cph_muni, fill = NA, color = "grey50", size = 0.05) +
  geom_sf(data = grids100m$data_poly[[1]],
          fill = NA,
          color = "red",
          size = 0.05) + 
  my_theme_map() +
  labs(title = "B) Grid cells of 100m x 100m") +
  annotation_scale(location = "br", text_cex = 1) +
  annotation_north_arrow(location = "br",
                         pad_x = unit(1.30, "cm"),
                         pad_y = unit(0.65, "cm"),
                         which_north = "true",
                         height = unit(0.5, "cm"),
                         width = unit(0.5, "cm"),
                         style = north_arrow_orienteering(text_col = "white",
                                                          text_size = 1))

# Plot
p1 + p2

```

## Population density

We disaggregate therefore population data at municipality level to the grid cells of 100m x 100m (i.e. `r nrow(grids100m$data_points[[1]])` grid cells in the study area) using all residences for year-round living (*ENH_ANVEND_KODE*) (i.e. from the Building and Dwelling Register  - [BBR](https://teknik.bbr.dk/forside) as ancillary data. We also removed from the dataset those dwellings that are not actually used for residential purpose (i.e. *BOLIGTYPE_KODE $\neq$ E - Andet (bl.a. institutioner og erhverv)* or *BOLIGTYPE_KODE $\neq$ 5 - Sommer-/fritidsbolig*) and dwellings with a size lower than that 10 $m^2$. 

```{r load-BBR}

# Buildings for year round living
  res_codes <- tribble (~ENH_ANVEND_KODE, ~type,
                        110, "Farmhouse",
                        120, "Single-family house",
                        121, "Single-family house",
                        122, "Single-family house",
                        130, "Semi-detached house",
                        131, "Semi-detached house",
                        132, "Semi-detached house",
                        140, "Multi-storey",
                        150, "College", 
                        160, "Residential institution",
                        190, "Second building") %>%
    # Convert type to factor
    mutate(type = factor(type)) %>%
    # Convert to data.table
    as.data.table()

# Function for reading residential units in the study area from a BBR files (.csv):
f_res_units <- function(.file, .muni = study_area_codes) {
  fread(.file) %>% 
    # Select only Residential houses - Buildings for year-round living
    filter.(ENH_ANVEND_KODE %in% res_codes$ENH_ANVEND_KODE) %>%
    # Format KomKode and select municipalities of the study area
    mutate.(KomKode = paste0("0", KomKode)) %>% 
    filter.(KomKode %in% .muni) %>% 
    # Remove residential units not used for for residential purpose
    filter.(BOLIGTYPE_KODE != "E" | BOLIGTYPE_KODE != "5") %>% 
    # Remove tiny dwellings (area < 10 m2)
    filter.(BEBO_ARL >= 10) %>%
    # Add year of the BBR dataset 
    mutate.(year = parse_number(stringr::str_extract(.file, "_[0-9]+_"))) %>% 
    # Convert house price (KONTANT_KOEBESUM) to numeric and kDKK
    mutate.(price_kDKK = KONTANT_KOEBESUM / 1000,
            price_kDKK = as.numeric(price_kDKK)) %>% 
    # Drop unused factors levels
    droplevels()
}

# Load residential units (from .csv files)
# NOTE: you may need to change the path to your file on "OneDrive - Aalborg Universitet" 
  csv_files_path <- list.files(path = Sys.getenv("OneDrive_BBR_path"),
                               pattern = "*.csv",
                               full.names = TRUE)
  
  plan(multisession, workers = 7)
  res_units_read <- future_map_dfr(.x = csv_files_path, .f = f_res_units)
  plan("default")
  
# Clean dataset
  res_units <- res_units_read %>% 
    # Input empty cells (buildings with only one floor) in Etagebetegn as "st"
    mutate.(Etagebetegn = fifelse(Etagebetegn == "", "st", Etagebetegn)) %>% 
    # Etagebetegn as ordered factor
    mutate.(Etagebetegn = factor(Etagebetegn,
                                 c("k2", "kl", "st", seq(1, 36, 1)),
                                 ordered = TRUE)) %>% 
    # Group floor levels with 5 or more
    mutate.(floor_level = fct_other(Etagebetegn,
                                    drop = factor(seq(5, 36)),
                                    other_level = "5 or more")) %>% 
    # Add residential description (type) into the dataset
    left_join.(res_codes, by = "ENH_ANVEND_KODE") %>% 
    # Convert columns with codes (*_KODE) to character
    mutate.(across.(ends_with("KODE"), ~as.character(.))) 

# Convert to sf objects 
  res_units_sf <- res_units %>% 
    st_as_sf(coords = c("etrs89koordinat_ost",
                        "etrs89koordinat_nord"),
             crs = "EPSG:25832")
   
```

The population data at municipality level for the same period we have BBR data (2004-2019) have been retrieved from [Statistics Denmark](https://www.dst.dk/en). We have used the R-package [danstat](https://cran.r-project.org/web/packages/danstat/index.html) and the table [FT: Population figures from the censuses](https://www.statbank.dk/statbank5a/SelectVarVal/Define.asp?Maintable=FT&PLanguage=1).

```{r pop-data-cph}

id_table <- "FT"
dat_meta <- get_table_metadata(table_id = id_table,
                               variables_only = TRUE)

# Variables
  variables <- list(
    # Municipalities
    list(code = "HOVEDDELE", values = c("101","147")),
    # Population at the first day of the year (from 2004 to 2019)
    list(code = "Tid", values = seq(2004, 2019, 1))
    )

# Get data 
  pop_tot_muni <- get_data("FT", variables) %>%
    # Translate names into English 
    rename(muni_name = HOVEDDELE,
           date = TID, 
           pop_total = INDHOLD) %>%
    # Translate København
    mutate(muni_name = gsub("Copenhagen", "København", muni_name))

# Add spatial information
   pop_tot_muni_sf <- cph_muni %>% 
    left_join(pop_tot_muni, by = "muni_name")
   
```

The procedure for disaggregating the population data is therefore as follow:  

1. Calculate the occupancy rate (*OR*) for the residential units of each municipality (*j*): $$OR_{j} = \frac{pop_{j}}{N_{j}}$$

1. Make grid cells of 100m x 100m over the study area, and select only the grids with residential units 

1. Detect to what municipality (*j*) belong each grid (*i*) (Note that one grids may be in more that one municipality)

1. Calculate the number of dwellings per grid and municipality ($N_{ij}$)

1. Estimate the population in each grid (*i*) base the occupancy rate by parish (*j*): $$pop_{gi} = \sum_{j = 1}^{n}(OR_{j} \cdot N_{ij})$$

1. Population density: $$PD_{i} = \frac{pop_{i}}{A_{i}}$$

We can therefore estimate the population density at the first day of the year by grid cells of 100m x 100m (figure \@ref(fig:fig-pop-density)).

```{r fig-pop-density-aux-function}

#' Aux. function for calculate the population density in the grids created by f_grids 
#' @param .year       Year of the analysis (e.g. 2019)

f_pd_grids <- function(.year) {

  # Population at the first day of the year (.year) in each municipality
  pop_year <- pop_tot_muni_sf %>%
    # select the data at the first day of the year 
    filter(date == .year)  
  
  # Dwellings (points) of the selected year 
  BBR_year <- res_units_sf %>% 
    filter(year == .year) %>% 
    select(geometry)
  
  # Calculate occupancy rate ("OR") of the residential units in each municipality in a year
  OR <- pop_year %>%  
    # number of units per municipality
    mutate(n_units = st_intersects(., BBR_year) %>%
             map(., ~length(.)) %>%
             unlist()) %>% 
    # mean population per unit in each parish
    mutate(across(starts_with("pop"), ~ . / n_units)) %>% 
    # output as table 
    as_tibble() %>% 
    select(-n_units, -geometry)
  
  # Get only the grids with residential units on them
  gru <- grids100m$data_poly[[1]] %>% 
    # Number of residential units per grid
    mutate(n_units = st_intersects(., BBR_year) %>%
             map(., ~length(.)) %>%
             unlist()) %>% 
    # Get grids with residential buildings
    filter(n_units > 0) %>% 
    # Add date
    mutate(date = .year)
  
  # Detect to what area belong the grid
  gru_muni <- gru %>% 
    st_intersection(., cph_muni) %>% 
    # Remove parish area
    select(-area_km2, -muni_name) 
  
  # Population density by grids
  pop_grids <- gru_muni %>% 
    # convert to table
    as_tibble() %>% 
    select(-geometry) %>% 
    # Merge OR per area  
    left_join(OR, by = "muni_id") %>% 
    # recalculate population by grid 
    mutate(across(starts_with("pop"), ~ . * n_units)) %>% 
    # sum population of each municipality of the grid
    group_by(grid_ID) %>% 
    summarise(pop_total = sum(pop_total),
              n_units = sum(n_units)) %>% 
    ungroup() 
  
  # Add spatial information
  gru %>% 
    select(-n_units) %>% 
    left_join(pop_grids, by = "grid_ID") %>% 
    # Area of the grid
    mutate(area_km2 = as.numeric(units::set_units(st_area(.), km^2))) %>% 
    # Population density (pop/area)
    mutate(across(starts_with("pop_"), ~ . / (1000 * area_km2))) %>% 
    rename_with(~paste(.x, "km2", sep = "_"), starts_with("pop")) %>% 
    # Remove polygons with 0 population
    filter(pop_total_km2 > 0)
  
 }

```

```{r pop-density-cph}

# List of years with BBR data (2004-2019) and population data (2008-2019)
years_pd <- seq(2004, 2019, 1) 

# Increase future maximum allowed size of global variables
options(future.globals.maxSize = 8000 * 1024^2)

# Population density 
plan(multisession, workers = 7)
pop_g100m <- future_map_dfr(.x = years_pd, .f = f_pd_grids) 
plan("default")

# Default value
options(future.globals.maxSize = 500 * 1024^2)

```

```{r fig-pop-density-cph, animation.hook="gifski", fig.cap = "Population density by grid cells of 100m x 100m"}

# Add breaks
brks <- c(min(pop_g100m$pop_total_km2),
          1, 2, 3, 4, 5, 10, 15, 20, 30,
          ceiling(max(pop_g100m$pop_total_km2)))

pop_g100m <- pop_g100m %>%
  mutate(pop_total_km2_cut = cut(pop_total_km2,
                                 breaks = brks,
                                 include.lowest = TRUE))  

# Aux. function for plotting
f_plot <- function(df) {
  ggplot() +
    geom_sf(data = study_area, fill = "grey") +
    geom_sf(data = df, 
            aes(fill = pop_total_km2),
            color = NA) +
    # geom_sf(data = cph_muni, fill = NA, color = "grey50", size = 0.1) +
    scale_fill_viridis(name = TeX("$\\overset{\\textbf{Population}}{(x1000/km^2)}$"),
                       option = "turbo",
                       limits = c(0.2, 174),
                       trans = "log10") +
    my_theme_map() +
    labs(x = "",
         y = "",
         caption = paste("Date", df$date, sep = ": "))
}

# Plot each year
plan(multisession, workers = 7)
plts <- pop_g100m %>% 
  group_split(date) %>%  
  future_map(., .f = f_plot, .options = furrr_options(seed = 123))
plan("default")

# Animation
for(i in seq_along(years_pd)) { print(plts[[i]]) }

```

```{r fig-pop-density-cph-b, fig.cap = "Population density by grid cells of 100m x 100m"}

ggplot() +
  geom_sf(data = study_area, fill = "grey") +
  geom_sf(data = pop_g100m, 
          aes(fill = pop_total_km2),
          color = NA) +
  scale_fill_viridis(name = TeX("$\\overset{\\textbf{Population}}{(x1000/km^2)}$"),
                     option = "turbo",
                     trans = "log10") +
  my_theme_map() +
  labs(x = "",
       y = "") +
  facet_wrap(~date)

```

