# Housing prices at municipality level {#house-prices-LAUs}

We load all residences for year-round living (*ENH_ANVEND_KODE*) from the Building and Dwelling Register ([BBR](https://teknik.bbr.dk/forside)), and then we estimate the mean and median prices in each municipality. 

- 110           = Farmhouse for agricultural property.  
- 120, 121, 122 = Detached single-family house (detached house). 
- 130, 131, 132 = Townhouse, chain, or semi-detached house (vertical separation between the units).
- 140           = Multi-storey residential building (multi-family house, including two-family house (horizontal separation between the units).
- 150           = College.
- 160           = Residential building for residential institution.
- 190           = Second building for year-round living.


We selected from the main residential buildings those that are on the ordinary free trade (*OVERDRAGELSES_KODE == 1 - Almindelig frit salg*) or public sales  (*OVERDRAGELSES_KODE == 3 - Auktion*). Furthermore, we filtered those dwellings that are actually used for residential purpose (i.e. *BOLIGTYPE_KODE $\neq$ E - Andet (bl.a. institutioner og erhverv)* or *BOLIGTYPE_KODE $\neq$ 5 - Sommer-/fritidsbolig*). Dwelling with a size lower that 10 $m^2$ were removed from the analysis. *Colleges* and *Residential building for residential institution*  were also excluded from the data analysis since they are a special type of buildings dedicated to students residences and housing for institutions, respectively, and they are mainly outside of the free marked.

```{r res-units-DK}

# Residential buildings for year round living
res_codes <- tribble (~ENH_ANVEND_KODE, ~type,
                      110, "Farmhouse",
                      120, "Single-family house",
                      121, "Single-family house",
                      122, "Single-family house",
                      130, "Semi-detached house",
                      131, "Semi-detached house",
                      132, "Semi-detached house",
                      140, "Multi-storey",
                      190, "Second building") %>%
  # Convert type to factor
  mutate(type = factor(type)) %>%
  # Convert to data.table
  as.data.table()

# Aux. function
f_average_price_muni <- function(.file) {
  fread(.file) %>% 
    # Select only Residential houses - Buildings for year-round living
    filter.(ENH_ANVEND_KODE %in% res_codes$ENH_ANVEND_KODE) %>%
    # ordinary free trade or auction
    filter.(OVERDRAGELSES_KODE == "1" | OVERDRAGELSES_KODE == "3") %>% 
    # Remove residential units not used for for residential purpose
    filter.(BOLIGTYPE_KODE != "E" | BOLIGTYPE_KODE != "5") %>% 
    # Remove tiny dwellings (area < 10 m2)
    filter.(BEBO_ARL >= 10) %>%
    # Add year of the BBR dataset 
    mutate.(year = parse_number(stringr::str_extract(.file, "_[0-9]+_"))) %>% 
    # Convert house price (KONTANT_KOEBESUM) to numeric and kDKK
    mutate.(price_kDKK = KONTANT_KOEBESUM / 1000,
            price_kDKK = as.numeric(price_kDKK)) %>% 
    # Drop unused factors levels
    droplevels()
}

# Read in parallel 
csv_files_path <- list.files(path = Sys.getenv("OneDrive_BBR_path"),
                             pattern = "*.csv",
                             full.names = TRUE)

plan(multisession, workers = 7)
res_prices_read <- future_map_dfr(.x = csv_files_path,
                             .f = f_average_price_muni) %>% 
  # rename KomKode
  mutate.(KomKode = paste0("0", KomKode)) %>% 
  rename.(muni_id = KomKode)
plan("default")

```

There are some unrealistic house prices in *Kerteminde* and *Ringsted* municipalities. In this sense, there are 499 houses with a value of 267000000 DKK in *Kerteminde* and 310 with a value of 375000000 DKK in *Ringsted*. We may expect few houses with very high prices in this areas but these numbers seem unlikely for small towns and oly for one year (2008 and 2011, respectively). Therefore, we have removed them from the datasets assuming there was a mistake in the data imputation and we do not have other criteria for assigning a realistic value. We also remove  properties with no value (*KONTANT_KOEBESUM* = 0; there were 18,801 of 1,494,495 data)

```{r remove-unrealistic-values}

res_prices <- res_prices_read %>%
  filter(!(muni_id == "0440" & year == 2008 & KONTANT_KOEBESUM == 267000000)) %>%
  filter(!(muni_id == "0329" & year == 2011 & KONTANT_KOEBESUM == 375000000)) %>% 
  filter(KONTANT_KOEBESUM > 0)

```

We adjusted the housing prices to 2019 prices. In this regard, we take into  account the inflation and, therefore, prices from different years can be compared (@Valtersdorf2020). The adjusted price is obtained as follow:

$$ Pice_{2019} = Price_{i} \cdot \frac{Index_{2019}}{Index_{i}}$$

Where, $Price_{2019}$ is the adjusted housing price for 2019, $Price_{i}$ is the price of the respective year $i$, and $Index_{i}$ and $Index_{2019}$ are the price indexes for the origin year ${i}$ and $2019$, respectively. The indexes have been obtained from Statistic Denmark; i.e. table [EJ6: Price index for sales property (2006=100) by category of real property and unit](https://www.statistikbanken.dk/statbank5a/SelectVarVal/Define.asp?MainTable=EJ6&PLanguage=1&PXSId=0&wsid=cftree) (Table \@ref(tab:price-index-DK)). 

```{r price-index-DK}

# DST table (PRIS112 - Consumer price index (2015=100) by main figures and time)
  id_table <- "EJ6"
  dat_meta <- get_table_metadata(table_id = id_table, variables_only = TRUE)
  
# Values to retrieve
  variables <- list(
    # Category of real property (i.e. One-family houses)
    list(code = "EJENDOMSKATE", values = "0111"),
    # Index
    list(code = "TAL", values = "100"),
    # From 2004 to 2019
    list(code = "Tid", values = seq(2004, 2019, 1))
    )

# Get index
  price_index <- get_data("EJ6", variables) %>% 
    select(-TAL) %>% 
    # Translate into English 
    rename(index_desc = EJENDOMSKATE,
           index_year = TID, 
           index_value = INDHOLD) %>% 
    # Convert index tu numeric
    mutate(index_value = index_value/10) 
    
# Print price index table  
  price_index %>% 
    # Wide format
    pivot_wider(names_from = index_year, 
                values_from = index_value) %>% 
    kbl(caption = "Price index for sale properties in Denmark") %>% 
    kable_paper() %>% 
    scroll_box(width = "100%")
  
```

```{r adj-price-DK}

# Adjust housing prices
price_index_2019 <- price_index %>% 
  filter(index_year == 2019) %>% 

# Add adjusted prices 
res_prices <- res_prices %>% 
  # Add price index by year and category
  left_join.(price_index, by = c("year" = "index_year")) %>% 
  # Add columns with index_2019
  mutate.(index_2019 = price_index_2019$index_2019) %>% 
  # Calculate prices 2019
  mutate.(price_2019_kDKK = (price_kDKK * index_2019 / index_value)) %>% 
  # Price per m2
  mutate(price_2019_kDKK_m2 = price_2019_kDKK / BEBO_ARL)
 
```

Prices change (median) from 2008 to 2019.
```{r median-prices-muni}

# Summary statistics of 2019 adjusted housing prices by municipalities
res_prices_muni <- res_prices %>%
  # 2008  - year we start to have population data 
  filter.(year >= 2008) %>% 
  # Estimate mean and median values (house prices)
  summarise.(num_res_units = n(), 
             median_2019_kDKK_m2 = median(price_2019_kDKK_m2),
             .by = c(muni_id, year)) %>% 
  # Price change with 2008 as baseline
  arrange.(muni_id, year) %>% 
  mutate.(
    median_2019_kDKK_m2_dif = 
      100 * (median_2019_kDKK_m2 - first(median_2019_kDKK_m2)) / first(median_2019_kDKK_m2),
    .by = muni_id)
  
# Link house prices (at the first day of the year) with population.  
dk_muni_prices <- dk_muni_pop %>% 
  as_tibble() %>% 
  # Population at the first day of the year
  separate(date, c("year", "month", "day"), sep = "-") %>% 
  filter(month == "01") %>%
  mutate(year = as.numeric(year)) %>% 
  # get only population data by municipality
  select(muni_id, muni_name, area_km2, year, ancestry, pop, pop_km2, geometry) %>%
  # wide format
  pivot_wider(names_from = ancestry,
              values_from = c(pop, pop_km2)) %>% 
  # Add spatial residential prices 
  left_join(res_prices_muni, by = c("muni_id", "year")) %>% 
  st_sf() %>% 
  # Remove Christiansø (there is no residential houses)
  filter(muni_id != "0411")

```

```{r fig-prices-LAUs, fig.cap = "Adjusted 2019 housing prices variation from 2008 to 2019"}

dk_muni_prices %>% 
  filter(year == 2019) %>% 
  ggplot() +
    geom_sf(aes(fill = median_2019_kDKK_m2_dif),
            color = "grey", size = 0.05) +
    scale_fill_gradient2(name = "Change [%]",
                         low = "blue",
                         mid = "white",
                         high = "red",
                         midpoint = 0) +
    theme_void()

```

Relationship between population density and house prices (median values) at municipality level.

```{r fig-prices-pop-a,  fig.height=8, fig.width=10, fig.cap = "Association between population density and median housing prices at municipality level"}

# Plot
ggplot() +
  geom_point(data = dk_muni_prices,
             aes(x = pop_km2_Total,
                 y = median_2019_kDKK_m2,
                 colour = muni_name,
                 shape = factor(year))) +
  scale_shape_manual(values = 1:nlevels(factor(dk_muni_prices$year))) +
  scale_x_log10() +
  geom_mark_ellipse(data = dk_muni_prices_lbls,
                    aes(x = pop_km2_Total,
                        y = median_2019_kDKK_m2,
                        colour = muni_name,
                        label = muni_name),
                    label.fill = NA) +
  labs(title = "",
       x = "Population density [ppl/km2]",
       y = "2019 adjusted median values [kDKK/m2]",
       colour = "Municipality",
       shape = "Year") +
  theme_bw() +
  theme(legend.position = "bottom") +
  guides(shape = guide_legend(ncol = 2, title.position = "top"),
         colour = guide_legend(ncol = 7, title.position = "top")) +
  geom_smooth(data = dk_muni_prices,
              aes(x = pop_km2_Total,
                  y = median_2019_kDKK_m2),
              method = "lm", 
              se = F,
              colour = "blue") 

```

Plots variation over time (adjusted prices). NOTE: Zoom up to 80% change

```{r fig-price-change-muni, fig.height=25, fig.width=10, fig.cap = "Adjusted median house price (kDKK/m2) change from 2008 (municipalities shorted by population density in 2019)"}

# Short municipalities by total population density in 2008-Q1
dk_muni_prices <- dk_muni_prices %>%
  mutate(muni_name = fct_reorder2(muni_name, year, pop_km2_Total,
                                  .fun = last2))

# Plot
ggplot(data = dk_muni_prices,
            aes(x = year,
                y = median_2019_kDKK_m2_dif)) +
  geom_point() +
  geom_smooth(method = "loess",
              span = 0.75,
              se = FALSE) +
  geom_hline(yintercept = 0, linetype="dashed", color = "grey", size = 0.5) +
  facet_wrap(~muni_name, ncol = 6) +
  theme_bw() +
  theme(legend.position = "bottom",
        axis.text = element_text(size = 10),
        axis.title = element_text(size = 14),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        title = element_text(size = 14)) +
  scale_y_continuous(name = "[%]") +
  coord_cartesian(ylim = c(-80, 80)) +
  scale_x_continuous(name = "",
                     breaks = seq(2008, 2019, 3))


```


```{r fig-prices-pop-b,  fig.height=10, fig.width=10, fig.cap = "Association between population density and median housing prices at municipality level"}

# Big cities/urban areas
  lbls <- c("København", "Aarhus", "Odense", "Aalborg", "Frederiksberg",
            "Tønder", "Læsø" )
  dk_muni_prices_lbls <- dk_muni_prices %>% 
    filter(muni_name %in% lbls) 
  
# Plot
  ggplot(data = dk_muni_prices,
         aes(x = pop_km2_Total,
             y = median_2019_kDKK_m2)) +
    geom_point() +
    geom_smooth(method = "lm") +
    scale_x_log10() +
    labs(title = "",
         x = "Population density [ppl/km2]",
         y = "2019 adjusted median values [kDKK/m2]") +
    ggrepel::geom_text_repel(data = dk_muni_prices_lbls, 
                             aes(label = muni_name),
                             nudge_y = 20,
                             max.iter = 3e3) +
    facet_wrap(~year) +
    theme_bw()

```
