# Open Street Map {#OSM-features}

The location information has been obtained from [Open Street Map](https://wiki.openstreetmap.org/wiki/Map_features). We have selected services inside a buffer of approx. 2 km around the study area (Figure \@ref(fig:fig-osm-poly) and \@ref(fig:fig-osm-points)), assuming that people may use then (e.g. parks, parking areas) outside Copenhagen City but they are close enough to people houses. Then, we estimate the accessibility to each services (i.e. feature) in the centroid of the grid cell of 100m x 100m; i.e. Potential model [chapter 1](#hedonic-price-model). We selected two types of features; *i. areal features* and *ii. point features*; where the mass of the service ($M_j$) is the area of the polygon in $m^2$ or equal for all services (i.e. 1), respectively. On the other hand, the distance ($d_{ij}$ in $m^2$) is the distance between the centroid of the polygon (i.e. areas features) or the point (i.e. point feature) and the centroid of the grid cells. 

## Features

### Areal features

The selected features and their expected effect on housing prices are reported in Table \@ref(tab:tbl-osm-poly). 

Hospital, and Schools (e.g. schools) may create traffic and noise. Therefore, they may have a negative effect on housing prices (@Gultekin2006). 

```{r tbl-osm-poly}

# List of OSM data
# Available features: https://wiki.openstreetmap.org/wiki/Map_features 

osm_poly_list <- tribble(
  ~feature, ~key, ~value, ~exp_effect,
  # Transportation   
  "Parking", "amenity", "parking", "positive",
  # Green areas
  "Park", "leisure",  c("park", "garden", "playground"), "positive",
  "Allotment", "landuse", "allotments", "positive",
  "Meadow", "landuse", "meadow", "positive", 
  # Recreation ground
  "Recreation", "landuse", "recreation_ground", "positive",
  # Education
  "School", "amenity", c("school", "kindergarten"), "negative",
  "University", "amenity", c("university", "college"), "positive", 
  # Healthcare
  "Hospital", "amenity", c("hospital"), "negative"
)

osm_poly_list <- osm_poly_list %>%
  unnest(c(value)) %>% 
  as.data.table() 

osm_poly_list %>%
  group_by(feature, key, exp_effect) %>%
  summarise(value = paste(value, collapse = ", ")) %>%
  relocate(value, .after = key) %>% 
  kbl(caption = "Expeced effect of polygon features on housing prices") %>% 
  kable_paper()

```

```{r osm-poly, cache = TRUE}

# Bounding of the study area (approx. buffer of 2km)
box <- c(xmin = 714906, xmax = 733618, ymin = 6166579, ymax = 6184092)
bbox <- st_bbox(box)
bbox_sf <- st_as_sfc(bbox) %>%  st_set_crs("EPSG:25832") 

# Get polygons 
osm_poly_dnld <- oe_get(place = "Copenhagen",
                        layer = "multipolygons",
                        extra_tags = c("building", "railway"),
                        provider = "bbbike") %>% 
  # Convert to data.table
  as.data.table() %>% 
  # Long format 
  melt(id.vars = c("osm_id", "osm_way_id", "type", "name", "geometry"),
       variable.name = "key",
       value.name = "value")

# Merge and convert to sf
osm_poly <- merge(osm_poly_list, osm_poly_dnld, by = c("key", "value")) %>% 
  # Convert to sf object
  st_sf() %>% 
  # Transform CRS
  st_transform("EPSG:25832") %>% 
  # Remove non-valid polygons (i.e. 2 polygons)
  mutate(valid = st_is_valid(.)) %>% 
  filter(!is.na(valid)) %>% 
  # From Meadow get only "Kalvebod Fælled" 
  filter(feature != "Meadow" |
         (key == "landuse" & value == "meadow" & name == "Kalvebod Fælled")) %>%
  mutate(feature = case_when(feature == "Meadow" ~ "Kalvebod_Fælled",
                             TRUE ~ feature)) %>% 
  # Mass of the service (area of the polygons)
  mutate(m_service = as.numeric(units::set_units(st_area(.), m^2))) %>% 
  # crop to Bounding of the study area %>% 
  st_crop(bbox_sf) %>% 
  # Nest by feature
  group_by(feature) %>% 
  nest() %>% 
  ungroup() %>% 
  rename(data_poly = data) %>% 
  # Get centroids  
  mutate(data_points = map(data_poly, st_centroid))

```

Spatial distribution (Figure \@ref(fig:fig-osm-poly))
```{r fig-osm-poly, fig.cap="Spatial distribution of OSM areal features"}

osm_poly %>%
  unnest(cols = c(data_poly)) %>% 
  st_sf() %>% 
  ggplot() +
  geom_sf(data = st_crop(dk_country, bbox_sf), fill = "grey95") +
  geom_sf(data = cph_parish, fill = "grey85", color = "grey50", size = 0.05) +
  geom_sf(fill = "#D55E00", col = NA) +
  labs(caption = "Source: Open Street Map") +
  theme_void() +
  facet_wrap(~feature)

```

### Point features

Point features and their expected effect on housing prices (Table \@ref(tab:tbl-osm-points)). 

```{r tbl-osm-points}
# Sustenance (points)

osm_points_list <- tribble(
  ~feature, ~key, ~value, ~exp_effect,
  # Sustenance
  "Sustenance", "amenity", c("bar", "biergarten", "cafe", "fast_food", "food_court",
                             "ice_cream", "pub", "restaurant"), "negative",
  # Shops
  "Shop", "shop", c("supermarket", "mall", "general", "dairy", "department_store",
                    "butcher", "seafood", "bakery", "convenience"), "positive",
  # Entertainment, Arts & Culture....
  "Entmt_pos", "amenity", c("cinema", "social_centre", "theatre",
                            "community_centre", "arts_centre", "public_bookcase"), "positive",
  "Entmt_neg", "amenity", c("brothel", "casino", "gambling",
                            "love_hotel", "nightclub", "stripclub", "swingerclub"), "negative",
  # Healthcare
  "Social_fac", "amenity",  "social_facility", "negative",
  # Public transport
  "Underground", "railway", "subway_entrance", "positive",
  "Bus_stop", "highway", "bus_stop", "positive",
  "Train_stop", "railway", "stop", "positive") %>%
  unnest(cols = c(value)) %>% 
  as.data.table()

osm_points_list %>% 
  group_by(feature, key, exp_effect) %>%
  summarise(value = paste(value, collapse = ", ")) %>% 
  relocate(value, .after = key) %>%
  kbl(caption = "Expeced effect of points features on housing prices") %>% 
  kable_paper()

```

```{r osm-points}

osm_points_dnld <- oe_get(place = "Copenhagen",
            layer = "points",
            extra_tags = c("amenity", "shop", "railway", "highway"),
            provider = "bbbike") %>%  
  # Convert to data.table
  as.data.table() %>% 
  # Long format 
  melt(id.vars = c("osm_id", "name", "address", "geometry"),
       variable.name = "key",
       value.name = "value")

osm_points <-  merge(osm_points_list, osm_points_dnld, by = c("key", "value")) %>% 
  # Convert to sf object
  st_sf() %>% 
  # Transform CRS
  st_transform("EPSG:25832") %>% 
  # crop to Bounding of the study area %>% 
  st_crop(bbox_sf) %>% 
  # Mass of the service (The same for all points = 1)
  mutate(m_service = 1)  %>% 
  # Nest by feature
  group_by(feature) %>% 
  nest() %>% 
  rename(data_points = data)  
  
```

Spatial distribution (Figure \@ref(fig:fig-osm-points))
```{r fig-osm-points, fig.cap="Spatial distribution of OSM point features"}

osm_points %>%
  unnest(cols = c(data_points)) %>% 
  st_sf() %>% 
  ggplot() +
  geom_sf(data = st_crop(dk_country, bbox_sf), fill = "grey95") +
  geom_sf(data = cph_parish, fill = "grey85", color = "grey50", size = 0.05) +
  geom_sf(colour = "#D55E00", size = 0.5) +
  labs(caption = "Source: Open Street Map") +
  theme_void() +
  facet_wrap(~feature)

```

## Potential model

We have used exponential functions with impedance factors of 2 and spans of 5000m, 500m, or 300m depending on the service we modelled (Table \@ref(tab:tbl-sif-pot-model)). 
```{r tbl-sif-pot-model}
# Based on: https://riatelab.github.io/potential/articles/potential.html

# Set Spatial Interaction Function (fun, span, beta)
sif <- tribble(   ~feature, ~fun,  ~span, ~beta,
         "Kalvebod_Fælled",  "e",   5000,     2,
              "Recreation",  "e",    500,     2,
              "Train_stop",  "e",    500,     2,
             "Underground",  "e",    500,     2,
                 "Parking",  "e",    300,     2,
                    "Park",  "e",    300,     2,
                  "School",  "e",    300,     2,
              "University",  "e",    300,     2,
                "Hospital",  "e",    300,     2,
               "Allotment",  "e",    300,     2,
              "Sustenance",  "e",    300,     2,
               "Entmt_pos",  "e",    300,     2,
               "Entmt_neg",  "e",    300,     2,
              "Social_fac",  "e",    300,     2,
                    "Shop",  "e",    300,     2,
                "Bus_stop",  "e",    300,     2
         )

sif %>% 
  group_by(fun, beta, span,) %>%
  summarise(feature = paste(feature, collapse = ", ")) %>% 
  relocate(feature, .after = span) %>% 
  kbl(caption = "Spatial interaction parameters") %>% 
  kable_paper()

```

The spatial interaction functions are therefero as follow (Figure \@ref(fig:spat-inter)).
```{r spat-inter, fig.width = 12, fig.height = 8, fig.cap="Spatial interaction functions"}

par(mfrow=c(2,2))

plot_inter(fun = "e",  span = 300, beta = 2)
plot_inter(fun = "e",  span = 500, beta = 2)
plot_inter(fun = "e",  span = 5000, beta = 2)

```

```{r merge-OMS-features}

# Merge all OMS features (as points)
OMS <- osm_poly %>%
  select(feature, data_points) %>% 
  bind_rows(osm_points) %>% 
  # Add SIF parameters
  left_join(sif, by = "feature")

```

```{r est-pot-model}

# Aux. function for calculating the potential
f_pot <- function(df, fun, span, beta) {
  mcpotential(x = df,
              y = grids100m$data_points[[1]],
              var = "m_service",
              fun = fun,
              span = span,
              beta = beta,
              limit = 5 * span)
  }

# Estimate potential (in percentage - relatively to the maximum)
pot_est <- list()
for(i in seq_along(OMS$data_points)) {
  # Values
  potential <- f_pot(OMS$data_points[[i]],
                     OMS$fun[i], 
                     OMS$span[i],
                     OMS$beta[i])
  # Percentage relatively to the maximum
  pot_est[[i]] <- 100 * potential / max(potential)
  # Add names
  names(pot_est)[[i]] <- OMS$feature[i] 
}

# Add values to the grid cells of 100 x 100 m (grids100m)
pot <- rbind(pot_est) %>%
  as_tibble() %>% 
  unnest(everything())

grids100m <- grids100m %>% 
  bind_cols(pot %>% nest(pot = everything()))

```

Potential model (Figure \@ref(fig:fig-pot-model))
```{r fig-pot-model, fig.cap="Potential model"}

# Plots
grids100m %>% 
  select(data_poly, pot) %>% 
  unnest(c(data_poly, pot)) %>% 
  pivot_longer(-c(grid_ID, geometry),
               names_to = "feature",
               values_to = "value") %>% 
  st_sf() %>% 
  ggplot() +
  geom_sf(aes(fill = value), col = NA) +
  scale_fill_gradient2(name = "[%]",
                         low = "blue",
                         mid = "white",
                         high = "red",
                         midpoint = 50) +
  geom_sf(data = cph_parish, fill = NA, color = "grey50", size = 0.05) +
  theme_void() +
  facet_wrap(~feature)
  
```
