# Distance to Central Business District {#dist-CBD}

The distance between a house and the Central Business District (CBD) has an important role in its price (@Chen2008). We add thereofre it as another feature in our model, with a possitive expected effec (i.e. the closer the house to the CBT, the higer the price). However, we do not measure the distance in km but the commuting time from the house to the CBD (i.e. "Indre By"). In this regard, we estimate the travel time (in public transport) from the centre of each grid cell of 100m x 100m to "Kongens Nytorv" (coordinates: 55.6805° N, 12.5860° E), and we associate the time to all dwellings at that grid. We have used used the  open source [OpenTripPlanner](https://www.opentripplanner.org/), and the r-package [opentripplanner](https://github.com/ropensci/opentripplanner) (@Morgan2019) for calculating the travel times by foot, bike, car, and public transport (Figure \@ref(fig:fig-travel-times)).  

```{r Kongens-Nytorv}

# Kongens Nytorv

KN <- tibble(name = "Kongens Nytorv", x = 12.5860, y = 55.6805) %>% 
  st_as_sf(coords = c("x", "y")) %>% 
  st_set_crs("EPSG:4326") 
  
```
 
```{r open-trip-planner-data, results='hide'}

# OTP folder structure
dir.create("otp")
dir.create("otp/graphs/default")

path_data <- file.path("./otp")

# Data

## Download Denmark GTFS from dk_gtfs <- "http://www.rejseplanen.info/labs/GTFS.zip"
## and save as "~./otp/graphs/default/gtfs.zip")

## Copenhagen pbf
cph_match <- oe_match("Copenhagen",
                      provider = "bbbike",
                      quiet = TRUE)
oe_download(cph_match$url,
            provider = "bbbike",
            download_directory = "./otp/graphs/default",
            quiet = TRUE)
file.rename(from = "./otp/graphs/default/bbbike_Copenhagen.osm.pbf",
            to = "./otp/graphs/default/osm.pbf")

## Get DEM

# Download elevation model (10m resolution)
zip_files <- c("DTM10_616_71_ASC_UTM32-EUREF89.ZIP",
               "DTM10_616_72_ASC_UTM32-EUREF89.ZIP",
               "DTM10_617_71_ASC_UTM32-EUREF89.ZIP",
               "DTM10_617_72_ASC_UTM32-EUREF89.ZIP",
               "DTM10_618_71_ASC_UTM32-EUREF89.ZIP",
               "DTM10_618_72_ASC_UTM32-EUREF89.ZIP")

f1 <- function(zip_file){ 
  dangeo_get_data(ftp_folder = "dhm_danmarks_hoejdemodel/DHM-HISTORIK/DHM-2007/DTM_10M",
                  zip_name   = zip_file)
}

map(zip_files, f1)

# Read files
asc_files_path <- list.files(path = loc_dir ,
                             pattern = "*.asc",
                             full.names = TRUE, 
                             recursive = TRUE)

# Read blocks
dtm10_blocks <- map(.x = asc_files_path, .f = read_stars)

# Set crs (UTM32-ETRS89)
dtm10_blocks <- map(dtm10_blocks, ~st_set_crs(., st_crs(study_area)))

# Build a mosaic 
dtm10 <- st_mosaic(dtm10_blocks[[1]],
                   dtm10_blocks[[2]],
                   dtm10_blocks[[3]],
                   dtm10_blocks[[4]],
                   dtm10_blocks[[5]],
                   dtm10_blocks[[6]])

# Intersect with study area
cph_dtm10 <- st_crop(dtm10, study_area)

# Warping the raster
newgrid <- study_area %>%
  st_transform(crs = "EPSG:4326") %>%
  st_bbox() %>%
  st_as_stars(dx = 0.000065, dy = 0.000065) # approx. 10m

cph_dtm10_WGS84 <- cph_dtm10 %>% st_warp(newgrid)

# Export to otp folder 
write_stars(cph_dtm10_WGS84, "otp/graphs/default/dem.tif")

```

```{r open-trip-planner}

# Open Trip Planner

## Downloading OTP
path_otp <- otp_dl_jar()

## Building an OTP Graph
log1 <- otp_build_graph(otp = path_otp, dir = path_data, memory = 10240)

## Start OTP
log2 <- otp_setup(otp = path_otp, dir = path_data)

## Connect R to OTP
otpcon <- otp_connect()

```

```{r travel-times}

# Origin points
op <- grids100m$data_points[[1]] %>%
  st_transform(crs = "EPSG:4326")

# Calculate travel times (i.e. walk, bike, car, and public transport)
f_tt <- function(feature){
  otp_plan(otpcon = otpcon,
           fromPlace = op,
           toPlace = c(12.5860, 55.6805),
           get_geometry = FALSE,
           mode = feature,
           ncores = as.numeric(availableCores() - 1)) %>% 
    select(fromPlace, toPlace, duration)
}

dur_features <- list(walk = "WALK",
                     bike = "BICYCLE",
                     car = "CAR",
                     transit = c("WALK", "TRANSIT"))

tt <- map(dur_features, f_tt)

# Intersect with grid cells of 100 x 100 m
f1 <- . %>% 
  as_tibble() %>% 
  mutate(duration = duration/60) %>% 
  separate(fromPlace, c("Y", "X"), sep = ",") %>% 
  st_as_sf(coords = c("X", "Y")) %>% 
  st_set_crs("EPSG:4326") %>% 
  st_transform(crs = "EPSG:25832") %>% 
  st_intersection(grids100m$data_poly[[1]])

tt_grids_cent <- tt %>% map(~f1(.))

# Add to grids polygons
f2 <- . %>% 
  st_drop_geometry() %>% 
  left_join(as_tibble(grids100m$data_poly[[1]]), by = "grid_ID") %>% 
  st_sf()

tt_grids_poly <- tt_grids_cent %>%  map(~f2(.))

f3 <- function(feature){
  tt_grids_poly %>% 
    pluck(feature) %>% 
    mutate(feature = feature)
}

tt_grids_poly <- names(tt_grids_poly) %>% map(~f3(.))

# Merge grids
tt_cph_grids100m <- bind_rows(tt_grids_poly) %>% 
  mutate(feature = factor(feature, levels = c("walk", "bike", "car", "transit")))

saveRDS(tt_cph_grids100m, file = "travel_times_grids100m_otp.rds")

```

```{r stop-otp, results='hide'}

# Stop OTP
otp_stop(warn = FALSE)

```

```{r fig-travel-times, fig.cap="Travel times to Kongens Nytorv (black point)"}

tt_cph_grids100m <- readRDS(file = "travel_times_grids100m_otp.rds")

myPalette <- RColorBrewer::brewer.pal(8, "RdYlBu")

ggplot() +
  geom_sf(data = tt_cph_grids100m, 
          aes(fill = cut(duration,
                         breaks = c(0, 10, 20, 30, 40, 50, 60, 120,
                                    ceiling(max(duration))),
                         include.lowest = TRUE)),
          col = NA) +
  scale_fill_manual(name = "min", values = rev(myPalette)) +
  geom_sf(data = cph_parish, fill = NA, color = "grey50", size = 0.05) +
  geom_sf(data = KN, color = "black", shape = 16, size = 1.5) +
  theme_void() +
  facet_wrap(~feature)

```

