# Open Street Map 

We get location information from [Open Street Map](https://wiki.openstreetmap.org/wiki/Map_features). We select services inside a buffer of approx. 2 km around the study area (Figure \@ref(fig:loc-box)), assuming that people may use then (e.g. parks, parking areas) outside Copenhagen City but they are close enough to they houses. 

```{r loc-box, cache=TRUE}

# Area for OSM data: Copenhagen
CPH_osm_bb <- getbb("Copenhagen")

# Bounding of the study area (approx. buffer of 2km)
box <- c(xmin = 714906,
         xmax = 733618,
         ymin = 6166579,
         ymax = 6184092)
bbox <- st_bbox(box)

bbox_sf <- st_as_sfc(bbox) %>% 
  st_set_crs("EPSG:25832") 

```

## Parking areas

Parking areas are tagged as `key = amenity` and `value = parking` (Figure \@ref(fig:osm-parking)). 
```{r osm-parking, fig.cap="Parking areas"}

# Get car packs from OSM
parking_osm <- CPH_osm_bb %>%
  opq() %>%
  add_osm_feature(key = "amenity", value = c("parking")) %>%
  osmdata_sf() 

# Clean dataset
parking <- parking_osm$osm_polygons %>%
  # Transform CRS
  st_transform(crs = "EPSG:25832") %>% 
  select(osm_id, amenity) %>% 
  # Calculate the area of the polygons
  mutate(area_m2 = as.numeric(units::set_units(st_area(.), m^2))) %>% 
  # Crop to area of influence
  st_crop(bbox_sf)

# Plot
ggplot() +
  geom_sf(data = st_crop(dk_country, bbox_sf),
          fill = "grey95") +
  geom_sf(data = cph_parish,
          fill = "grey85",
          color = "grey50",
          size = 0.05) +
  geom_sf(data = parking, fill = "#D55E00", col = NA) +
  my_theme_map() +
  labs(caption = "Source: Open Street Map")

```

Potential model
```{r set-pot-model-parking, cache=TRUE, dependson="loc-box"}

# Based on: https://riatelab.github.io/potential/articles/potential.html

# Points to estimate the potential 
grids100m_cent <- st_centroid(grids100m)

# Calculate centroid of the parking area
parking_cent <- st_centroid(parking)

# Distance matrix
d <- create_matrix(x = parking_cent, y = grids100m_cent)

```

Spatial interaction function (Figure \@ref(fig:spat-inter-parking)).
```{r spat-inter-parking, fig.cap="Spatial interaction function"}

span <- 300
beta <- 2
plot_inter(fun = "e",
           span = span,
           beta = beta)

```

Calculate the potential (Figure \@ref(fig:pot-model-parking)).
```{r pot-model-parking, fig.cap="Potential of parking areas"}

# Calculate the potential
grids100m_cent$pot <- potential(x = parking_cent,
                                y = grids100m_cent,
                                d = d,
                                var = "area_m2",
                                fun = "e",
                                span = span,
                                beta = beta)

# Potential relatively to its maximum
grids100m_cent$pot_perc <- 100 * grids100m_cent$pot / max(grids100m_cent$pot)

# Add info to grids100m
grids100m_cent <- grids100m_cent %>%
  as_tibble() %>%
  select(-geometry, int)

parking_pot <- grids100m %>% 
  as_tibble() %>% 
  left_join(grids100m_cent, by = "grid_ID") %>% 
  st_sf()

# Plot
ggplot() +
  geom_sf(data = parking_pot, aes(fill = pot_perc), col = NA) +
  scale_fill_viridis_c(name = "Perc [%]", option = "plasma") +
  geom_sf(data = cph_parish, fill = NA, color = "grey50", size = 0.05) +
  my_theme_map() 

```

## Parks

"A park is an area of open space for recreational use, usually designed and in semi-natural state with grassy areas, trees and bushes - OMS"  (Figure \@ref(fig:green-areas-osm)).
```{r green-areas-osm, fig.cap="Green areas"}

# Green areas 
  CPH_parks <- CPH_osm_bb %>%
    opq() %>%
    add_osm_feature(key = "leisure", value = c("park",
                                               "common",
                                               "garden",
                                               "firepit",
                                               "nature_reserve")) %>%
    osmdata_sf() 
  
# Clean
  parks <- select(CPH_parks$osm_polygons, osm_id, name) %>% 
    # Transform CRS
    st_transform(crs = "EPSG:25832") %>% 
    # Encoding names (Danish)
    mutate(name = iconv(name, "UTF-8")) %>% 
    # Calculate the area of the polygons
    mutate(area_m2 = as.numeric(units::set_units(st_area(.), m^2))) %>% 
    # Crop to area of influence
    st_crop(bbox_sf) 
    
# Plot
ggplot() +
  geom_sf(data =st_crop(dk_country, bbox_sf),
           fill = "grey95") +
  geom_sf(data = cph_parish,
          fill = "grey85",
          color = "grey50",
          size = 0.05) +
  geom_sf(data = parks,
          colour = NA,
          fill = "palegreen3",
          alpha = .3) +
  my_theme_map()
  
```

Potential model
```{r set-pot-model-parks, cache=TRUE, dependson="loc-box"}

# Points to estimate the potential 
grids100m_cent <- st_centroid(grids100m)

# Calculate centroid of the parking area
parks_cent <- st_centroid(parks)

# Distance matrix
d <- create_matrix(x = parks_cent, y = grids100m_cent)

```

Spatial interaction function (Figure \@ref(fig:spat-inter-parks)).
```{r spat-inter-parks, fig.cap="Spatial interaction function"}

span <- 300
beta <- 2
plot_inter(fun = "e",
           span = span,
           beta = beta)

```

Calculate the potential (Figure \@ref(fig:pot-model-parks)).
```{r pot-model-parks, fig.cap="Potential of parking areas"}

# Calculate the potential
grids100m_cent$pot <- potential(x = parks_cent,
                                y = grids100m_cent,
                                d = d,
                                var = "area_m2",
                                fun = "e",
                                span = span,
                                beta = beta)

# Potential relatively to its maximum
grids100m_cent$pot_perc <- 100 * grids100m_cent$pot / max(grids100m_cent$pot)

# Add info to grids100m
grids100m_cent <- grids100m_cent %>%
  as_tibble() %>%
  select(-geometry, int)

parks_pot <- grids100m %>% 
  as_tibble() %>% 
  left_join(grids100m_cent, by = "grid_ID") %>% 
  st_sf()

# Plot
ggplot() +
  geom_sf(data = parks_pot, aes(fill = pot_perc), col = NA) +
  scale_fill_viridis_c(name = "Perc [%]", option = "plasma") +
  geom_sf(data = cph_parish, fill = NA, color = "grey50", size = 0.05) +
  my_theme_map() 

```

# Forest

```{r}

CPH_forest <- CPH_osm_bb %>%
    opq() %>%
    add_osm_feature(key = "landuse", value = c("meadow", "allotments", "forest")) %>%
    osmdata_sf() 
  
  CPH_scrub <- CPH_osm_bb %>%
    opq() %>% 
    add_osm_feature(key = "natural", value = "scrub") %>% 
    osmdata_sf() 
  
  # Land_use: recreation_ground
  
```

