# Residential units saled as free scale {#EDA-house-prices}

We focused our study in residential dwellings on the ordinary free trade (Table \@ref(tab:tbl-runits-clean)). 

NOTES: 

- There are large differences in the number of residential units between years, why??
- Study only one year for the moment (e.g. 2019)??
- Housing prices in different years -> adjust to 2019 prices (what index; Table \@ref(tab:price-index))?
- **KONTANT_KOEBESUM** = cash purchase price? Analyse it?? What about **KOEBESUM_BELOEB** = The purchase price agreed upon the sale of the property???
- **KONTANT_KOEBESUM** = 0 or (< 100000 DKK)? 
- **SKOEDE_DATO** = "The date on which the deed was signed" -> use as sale date?? Should we use this date (i.e. for the price index)? ...   
 - What about housing prices = 0?
 - Dwelling Area <= 0? 
 
```{r}

res_units_oft %>% 
  ggplot(aes(sample = log10(price2019_kDKK),  colour = factor(BBR_year))) +
  geom_qq() +
  #geom_qq_line() +
  facet_grid(~type) +
  theme_bw()

```


## Residential units by floor level

- KL - basement
- ST - ground floor
- 1  - 1st floor
- 2  - 2nd floor
- 3  - 3rd floor
- 4  - 4th floor
- 5 or more = etc.

```{r fig-runits-floor, fig.cap = "Residential units distribution by floor level"}

  res_units_oft %>% 
    group_by(floor_level, type) %>% 
    summarise(n = n()) %>%
    ungroup() %>% 
    # in percentage [%]   
    mutate(perc = 100 * n / sum(n)) %>% 
    # Reorder type levels for plotting multi-storey first
    mutate(type = factor(type)) %>%
    ggplot() +
    geom_bar(aes(y = floor_level, x = perc, fill = type), stat = "identity") +
    labs(y = "", 
         x = "Percentage [%]") +
    theme_bw() +
    theme(legend.position = "bottom",
          legend.title = element_blank()) +
    guides(fill = guide_legend(ncol = 2)) +
    scale_x_continuous(labels = scales::comma) +
    scale_fill_manual(values = c("#0072B2", "#D55E00", "#CC79A7"))
  
```


